---
title: "Analysis of Heart Disease"
author: "Elias Fedai"
date: "8/5/2019"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


```{r, warning=FALSE,message=FALSE}
library(ggplot2)
library(knitr)
library(mlr)
library(tidyverse)
library(GGally)
library(cowplot)
library(dplyr)
library(tidyr)
library(readr)
library(magrittr)
library(moments)
library(plyr)
library(caret)
library(caTools)
library(corrplot)
library(ggcorrplot)
library(mlbench)
library(pROC)
library(e1071)
clev <- read.csv(file.choose(),na = "?", stringsAsFactors = FALSE, header=F)
swiss <- read.csv(file.choose(),na = "?", stringsAsFactors = FALSE, header=F)
va <- read.csv(file.choose(),na = "?", stringsAsFactors = FALSE, header=F)
combined  <-rbind(clev,swiss,va)
names(combined) <- c('Age', 'Gender', 'CP', 'Trestbps', 'Chol', 'FBS', 'RestECG', 'Thalach', 'Exang', 'Oldpeak', 'Slope', 'CA','Thal', 'target')
summarizeColumns(combined) %>% knitr::kable( caption =  'Initial Summary of Data')
```

*Data needs to be cleaned.

```{r}

 
combined[, sapply(combined, is.character)] <- sapply(combined[, sapply(combined, is.character)], trimws)
 s = sum(is.na(combined))
 combinedclean <- na.omit(combined)
combinedclean$target[combinedclean$target == 2] <- 1
combinedclean$target[combinedclean$target == 3] <- 1
combinedclean$target[combinedclean$target == 4] <- 1
combinedclean$Disease <- factor(combinedclean$target, labels = c("No Disease", "Heart Disease"))
combinedclean$target <- as.factor(combinedclean$target)
summarizeColumns(combinedclean) %>% knitr::kable( caption = 'Summary of Data (clean)')
table(combinedclean$target) %>% kable(caption = 'Table of Heart Disease')
```


*There seems to be a lower count on heart disease cases than non heart disease cases.


```{r}
combinedclean$Gender[combinedclean$Gender == 0] <- 'Female'
combinedclean$Gender[combinedclean$Gender == 1] <- 'Male'
 combinedclean$Gender <- as.factor(combinedclean$Gender)
 combinedclean$CP[combinedclean$CP == 1] <- 'typical angina'
combinedclean$CP[combinedclean$CP == 2] <- 'atypical angina'
combinedclean$CP[combinedclean$CP == 3] <- 'non-anginal pain'
combinedclean$CP[combinedclean$CP == 4] <- 'asymptomatic'
combinedclean$CP <- as.factor(combinedclean$CP)
combinedclean$Exang[combinedclean$Exang == 0] <- 'no'
combinedclean$Exang[combinedclean$Exang == 1] <- 'yes'
combinedclean$Exang <- as.factor(combinedclean$Exang)
combinedclean$Thal[combinedclean$Thal == 3] <- 'normal'
combinedclean$Thal[combinedclean$Thal == 6] <- 'fixed defect'
combinedclean$Thal[combinedclean$Thal == 7] <- 'reversable defect'
combinedclean$Thal <- as.factor(combinedclean$Thal)
combinedclean$FBS[combinedclean$FBS == 0] <- 'false'
combinedclean$FBS[combinedclean$FBS == 1] <- 'true'
combinedclean$FBS <- as.factor(combinedclean$FBS)
combinedclean$RestECG[combinedclean$RestECG == 0] <- 'normal'
combinedclean$RestECG[combinedclean$RestECG == 1] <- 'ST-T abnorm'
combinedclean$RestECG[combinedclean$RestECG == 2] <- 'V.hypertrophy'
combinedclean$RestECG <- as.factor(combinedclean$RestECG)
combinedclean$Slope[combinedclean$Slope == 1] <- 'upsloping'
combinedclean$Slope[combinedclean$Slope == 2] <- 'flat'
combinedclean$Slope[combinedclean$Slope == 3] <- 'downsloping'
combinedclean$Slope <-as.factor(combinedclean$Slope)
```
### Cholesterol
```{r}
summary(combinedclean$Chol)
hist(combinedclean$Chol, main = "Histogram of Cholesterol", xlab = "Cholesterol(mg/dL)", border = "blue", col ="red", xlim=c(100,700), las=1, breaks = 7, probability = TRUE)
lines(density(combinedclean$Chol))
boxplot(combinedclean$Chol ~ combinedclean$Disease, main = "Cholestrol level", xlab = "Disease", ylab = "Cholestrol level(mg/dL)")
```

*Both heart and no disease display slightly elevated cholestrol levels.

### Gender
```{r}
summary(combinedclean$Gender)
mosaicplot(combinedclean$Gender ~ combinedclean$Disease, main= " Heart Disease by Gender", shade = FALSE,  color = TRUE, xlab = "Gender", ylab = "Heart Disease")
ggplot(combinedclean, aes(x= combinedclean$Gender, fill=combinedclean$Disease)) + geom_bar() + xlab("Gender") +   ylab("Gender Count") + ggtitle("Heart Disease by Gender") +  scale_fill_discrete(name = "Heart disease", labels = c("No", "Yes"))
```
*Graphs/tables display more males where used in this study also there is a higher frequency of males with heart disease.


### Thalassemia
```{r}
summary(combinedclean$Thal)

ggplot(combinedclean, aes(x=Thal, fill=Disease)) + geom_bar()  + ggtitle("Analysis of Thalasemia(Thal)") + scale_fill_discrete(name = "Heart Disease", labels = c("No", "Yes"))
```
*Reversable defect displays a higher measure for heart disease.


### Oldpeak
```{r}
summary(combinedclean$Oldpeak)

ggplot(combinedclean, aes(x = Oldpeak, fill = Disease)) +  xlab("ST depression induced by exercise relative to rest ") + ylab("Count") + geom_density(alpha=0.5) + ggtitle("ST depression and heart disease") + scale_fill_discrete(name = "Heart disease", labels = c("No", "Yes"))

boxplot(combinedclean$Oldpeak ~ combinedclean$Disease, main = "Oldpeak and Heart Disease", xlab = "Disease", ylab = "ST depression induced by exercise relative to rest")

ggplot(combinedclean, aes(x=Oldpeak, fill=Disease)) + geom_bar()  + ggtitle("ST depression induced by exercise(Oldpeak) relative to rest and Heart Disease") + scale_fill_discrete(name = "Heart Disease", labels = c("No", "Yes"))

```
*Visuals for ST levels show a heavy right skewness. The no disease displays slightly more outliers.

### Fasting blood sugar
```{r}
summary(combinedclean$FBS)
ggplot(combinedclean, aes(x=FBS, fill=Disease)) + geom_bar()  + ggtitle("Fasting Blood Sugar(FBS)") + scale_fill_discrete(name = "Heart Disease", labels = c("No disease", "Heart Disease"))

mosaicplot(combinedclean$FBS ~ combinedclean$Disease, main= "Heart Disease by Fasting Blood sugar(FBS)", shade = FALSE,  color = TRUE, xlab = "FBS", ylab = "Heart Disease")

```
*Visuals appear to show not much possible significance for fasting blood sugar.

### Resting Electrocardiographic
  
```{r}
summary(combinedclean$RestECG)

ggplot(combinedclean, aes(x=RestECG, fill=Disease)) + geom_bar()  + ggtitle("Resting Electrocardiographic results") + scale_fill_discrete(name = "Heart Disease", labels = c("No", "Yes"))


```
*V hypertrophy displayed a higher count for heart disease.



### Logistic regression

```{r}
set.seed(300)

ind1 <-sample(2, nrow(combinedclean), replace = T, prob = c(0.7, 0.3))
 training1 <- combinedclean[ind1 == 1,]
 test1 <- combinedclean[ind1 == 2,]
 
 model1 <- glm(target ~ Gender + Age + CP +  Trestbps + RestECG +  Thalach + Exang + Slope + CA + Thal +  Oldpeak + FBS, data = combinedclean, family = binomial)
summary(model1)
 
```
*Parameters that were seen to be insignificant will be removed.
  
  
```{r}
model1 <- glm(target ~ Gender + CP + CA + Trestbps + Thal + Oldpeak, data = training1, family = binomial)
model2 <- glm(target ~ Gender + CP + CA + Trestbps + Thal + Oldpeak, data = test1, family = binomial)
summary(model1)

p1 <- predict(model1, training1, type = "response")
 pred1 <- ifelse(p1>0.5, 1, 0)
 tab1 <- table(Predicted = pred1, Actual = training1$target)
 tab1
 1- sum(diag(tab1))/sum(tab1)
```
*Training model displayed 18% misclassification error. Which will give an Accuracy of 82%

```{r}
with(model1, pchisq(null.deviance - deviance,  df.null-df.residual, lower.tail = F))
```


*fit test p value indicates this model is significant.


```{r}
summary(model2)
p2 <- predict(model2, test1, type = "response")
pred2 <- ifelse(p2>0.5, 1, 0)
tab2 <- table(Predicted = pred2, Actual = test1$target)
tab2
 1- sum(diag(tab2))/sum(tab2)
```

*Test model displayed a 13.5% misclassification error. Which will give an Accuracy of 86.5%
```{r}
with(model2, pchisq(null.deviance - deviance,  df.null-df.residual, lower.tail = F))
```


*Fit test p value indicates this model is signicant.
 
```{r}
ind2 <-sample(2, nrow(combinedclean), replace = T, prob = c(0.7, 0.3))
 training2 <- combinedclean[ind2 == 1,]
 test2 <- combinedclean[ind2 == 2,]
 
 model3 <- glm(target ~  RestECG + Age, data = training2, family = binomial)
 model4 <- glm(target ~  RestECG + Age, data = test2, family = binomial)
summary(model3)
p3 <- predict(model3, training2, type = "response")
 pred3 <- ifelse(p3>0.5, 1, 0)
 tab3 <- table(Predicted = pred3, Actual = training2$target)
 tab3
  1- sum(diag(tab3))/sum(tab3)
```


*training model displayed 32.2% misclassifcation error.
 which will give an Accuracy of 67.8%
 
```{r}
with(model3, pchisq(null.deviance - deviance,  df.null-df.residual, lower.tail = F))
```


*Fit test p value indicates this model is signifcant.
 
```{r}
summary(model4)
  p4 <- predict(model4, test2, type = "response")
 pred4 <- ifelse(p4>0.5, 1, 0)
 tab4 <- table(Predicted = pred4, Actual = test2$target)
 tab4
  1- sum(diag(tab4))/sum(tab4)
```
*Test model displayed 45.6% misclassification error. Which will give an Accuracy of 54.4%
 
 
```{r}
with(model4, pchisq(null.deviance - deviance,  df.null-df.residual, lower.tail = F))

```
*Fit test p value indicates this model is not significant.
  
